---
title: "Control Flow and Function Writing"
format: html
editor: visual
---

For Homework 3, we will be continuing the example in our last homework assignment. This file $hw2_list.rda$ is located in the working directory and is loaded in to work with in this assignment.

```{r}
load("hw2_list.rda")
```

# Task 1: Control Flow Practice

1. Suppose we want to characterize the post-treatment (or placebo) blood pressure measurement as optimal (≤ 120), borderline (120 < bp ≤ 130), and high (> 130). 

First, create a new column in each data frame from above called status. You can do this via

your_df$status <- character(20) #or 10 depending on number of observations

Note: You want to do this additional column to the data frames that are stored in your list not the original data frames you had in your environment. Although R uses smart ways to avoid using excess memory, it doesn’t overwrite the data frame stored in your list if you modify the original object.


```{r}
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```


2. For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.


```{r}
for(i in seq_len(nrow(bp_list$treatment)){
  if(bp_list$treatment$post_bp[i] > 130){
      bp_list$treatment$status[i] == "high"
  } else if (bp_list$treatment$post_bp[i]){
      bp_list$treatment$status[i] == "borderline"
  } else if (bp_list$treatment$post_bp[i]){
      bp_list$treatment$status[i] == "optimal"
  } else {
      bp_list$treatment$status[i] == "Error"
  }
}
```


```{r}
bp_list$treatment$status <- ifelse(bp_list$treatment$post_bp > 130, "high",
                           ifelse(bp_list$treatment$post_bp > 120, "borderline", "optimal"))
```




